Viscous Transport in Eroding Porous Media

### 2019-07-23 docs/transport/erosion_transport.tex
add figs in transport paper
add modified figs in transport paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 nearly touching bodies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies a Hagen-Poiseuille flow.  The background
flow is 
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right].
\end{align}
Initially, $U=1$, and then $U$ is updated at each time step so that the
pressure drop from $x=-2$ to $x=2$ remains constant.  Snapshots of the
bodies at four equispaced times are shown in
Figure~\ref{fig:Eroding20vort}.   The colormap is the vorticity which is
equivalent to the shear stress when restricted to the eroding bodies.
Therefore, erosion is fastest in regions where the magnitude of the
vorticity is largest.

@@

\begin{figure}[H]
\begin{center}
  \includegraphics[height=0.43\textwidth]{./figs/20b_dense1}
  \includegraphics[height=0.43\textwidth]{./figs/20b_dense101} \\
  \includegraphics[height=0.43\textwidth]{./figs/20b_dense201}
  \includegraphics[height=0.43\textwidth]{./figs/20b_dense301}
%\includegraphics[width = 0.42 \textwidth]{./figs/20b_dense1}
%\includegraphics[width = 0.507 \textwidth]{./figs/20b_dense101}\\
%\includegraphics[width = 0.42 \textwidth]{./figs/20b_dense201}
%\includegraphics[width = 0.507 \textwidth]{./figs/20b_dense301}
\caption{\label{fig:Eroding20vort} Erosion of 20 nearly touching bodies
  with the fixed pressure drop condition. The color is the vorticity of
  the fluid. The 4 snapshots are evenly spaced in time. We use $N_\iin =
  256$ discretization points on the body, a time-step of $\Dt=10^{-4}$,
  and smoothing parameters of $\eps= 15/256$ and $\sigma=10/256 $. In
  the forth frame, bodies 4, 12, and 17 have vanished completely, and
  body 19 has almost completely eroded.}
\end{center}
\end{figure}

In this example, we also observe that erosion creates a network of
channels from the inlet to outlet where the velocity and vorticity, and
therefore erosion rate, are much larger than the other regions.  These
channels can be further visualized by considering the streamlines at
several configurations of the eroding body.  In
Figure~\ref{fig:Eroding20tracer}, we simulate 200 tracers that are
initially to the left of the porous region and equispaced in $y$.  There
are two clear region where the tracers pass through the geometry
much faster than the other regions.  To illustrate the ability of the
Barycentric quadrature formula, streamlines very close to the eroding
bodies are illustrated in Figure~\ref{fig:Eroding20zoom}.

\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b60}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90}\\

\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b120}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b180}\\

\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b240}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270}
\caption{\label{fig:Eroding20tracer}The trajectories of 200 tracers in
  the fluid with 20 bodies. The 9 snapshots are evenly spaced in time.
  The tracers are evenly spread in the $y$-direction and released at
  $x=-1$.}
\end{center}
\end{figure}

@@

Since we use a Barycentric quadrature rule to approximate the velocity
of tracers and apply the fourth order Runge Kutta method to find the new
locations of tracers in each step, the tracers can be very close to but
not penetrate the bodies.  Those body-closed tracers are moving
perfectly so we are able to observe the details of the interaction of
bodies and fluid.  Although the direction of flow is from left to right,
the tracers will not only travel horizontally but also vertically.  In
the last three frames of Fig.~\ref{fig:Eroding20tracer}, some tracers
climb up from right to left around the body.  There are zoom windows of
the climbing tracers trajectories in Fig.~\ref{fig:Eroding20zoom} and
they are taken from the last three frames of
Fig.~\ref{fig:Eroding20tracer} with more tracers.  In
Fig.~\ref{fig:Eroding20zoom} the tracers around the center body first go
from the channel below the body, climb up to the left, then go to the
right and join another tracers in the channel above the body. 
}
\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.32\textwidth]{./figs/tracer_20b210_zoom}
\includegraphics[width = 0.32\textwidth]{./figs/tracer_20b240_zoom}
\includegraphics[width = 0.32\textwidth]{./figs/tracer_20b270_zoom}
\caption{\label{fig:Eroding20zoom}The zoom windows in the last three
  frames of Figure~\ref{fig:Eroding20tracer}. The trajectories of
  tracers around the bodies. Since we use a Barycentric quadrature rule
  and the fourth order Runge Kutta method, the tracers can be very close
  to but not penetrate the bodies.}
\end{center}
\end{figure}
{\color{red}
In Fig.~\ref{fig:Eroding20area}(a) we
measure the area occupied by bodies 
during the eroding. In the beginning, 62.32\% of area of $[-1,1] \times
[-1,1]$ is occupied by 20 bodies. Then the eroding speed is increasing
when more and more bodies are vanished in the fluid and 
the maximum eroding speed happens when 7.57\% of area of $[-1,1] \times
[-1,1]$ is occupied by 8 bodies. After that the speed drops rapidly to 0 when
all the bodies are vanished completely.

We record the flow rate $U$ quantifily in log scale
during the eroding in Fig.~\ref{fig:Eroding20area}(b). 
Since we simulate the erosion by applying fixed pressure drop 
the flow rate $U$ is small when there are many bodies in the fluid and 
when the area occupied by bodies is getting smaller $U$ is getting larger.
When all the bodies have vanished completely the flow rate $U$ is equal to 1.
In Fig.~\ref{fig:Eroding20area}(b), the flow rate increases exponentially 
with growth rate close to $\ln10^4$.

@@

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/porosity20dense}
\caption{}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/flow_rate20dense}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20area}  (a) The area fraction of 20 nearly
  touching eroding bodies and the eroding speed versus time. 
(b) The incoming flow rate of $U$ in log scale versus time. 
Under fixed pressure drop condition, $U$ will be small initially,
increase when bodies are eroding, and grow exponentially 
to 1 when there is no body in the fluid. }
\end{figure}

{\color{red}
In Fig.~\ref{fig:Eroding20tort}(c) we calculate the difference of 
$\tau$ at two neighboring tracers 
and draw the trajectories of neighboring tracers in 
Fig.~\ref{fig:Eroding20tort_traj}
if the difference of $\tau$ at the neighboring tracers is large. 
From Fig.~\ref{fig:Eroding20tort_traj},
 we are able to tell the neighboring trajectories will be very 
different if they are separated by bodies. In Fig.~\ref{fig:Eroding20tort_traj}
we collect the 10 largest difference of $\tau$ at the neighboring tracers and 
draw the pair of trajectories in the same color. For each pair of trajectories, 
they are splitted by the body at the left side of the body, then
one of them goes up and the other one goes down along the body 
until they rejoin with each other at the right side of body.  
Furthermore, the pair trajectories could be splitted by more than one body, 
for example in Fig.~\ref{fig:Eroding20tort_traj} the toppest red pair trajectories 
are splitted by two close bodies.
}

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc20_268}
\caption{}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_local20_268}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/diff_tort20_268}
\caption{}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/uminDumax20_268}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20tort} The tortuosity study in the fluid with 
20 nearly touching bodies when the porosity is 62.9 \%.
(a) The 
x-component velocity u(-1, y) with respect to its maximum velocity $u_{max}=2.9828
\times 10^{-3}$ at equal spreading tracers on the cross section x = -1. 
(b) The local tortuosity $\tau(y)$ on the cross section x = -1. 
(c) The difference of $\tau(y)$ at two neighboring tracers on the cross section x = -1.
(d) The ratio of the minimum to maximum speeds along the trajectories of tracers.}
\end{figure}

\begin{figure}[H]
\center
\includegraphics*[width =0.5\linewidth]{./figs/tort_diff_top10_268}
\caption{\label{fig:Eroding20tort_traj} The largest 10 difference of 
local tortuosity $\tau$ at two neighboring tracers in
 20 nearly touching bodies eroding when the porosity is 62.9 \% 
and the tortuosity is 1.0979. The same color of a pair of trajectories 
means they are from two neighboring tracers.}
\end{figure}

@@

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc100}
\caption{}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_local100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/diff_tort100}
\caption{}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/uminDumax100}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100tort} The tortuosity study in the fluid with 100 bodies 
when the porosity is 62.98 \%.
(a) The 
x-component velocity u(-1, y) with respect to its maximum velocity $u_{max}=3.9021
\times 10^{-4}$ at equal spreading tracers on the cross section x = -1. 
(b) The local tortuosity $\tau(y)$ on the cross section x = -1. 
(c) The difference of $\tau(y)$ at two neighboring tracers on the cross section x = -1.
(d) The ratio of the minimum to maximum speeds along the trajectories of tracers.}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 nearly touching bodies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies a Hagen-Poiseuille flow.  The background
flow is 
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right].
\end{align}
Initially, $U=1$, and then $U$ is updated at each time step so that the
pressure drop from $x=-2$ to $x=2$ remains constant.  
{\color{red}
We record the flow rate $U$ quantifily in log scale
during the eroding in Fig.~\ref{fig:Eroding20flowrate}. 
Since we simulate the erosion by applying fixed pressure drop 
the flow rate $U$ is small when there are many bodies in the fluid and 
when the area occupied by bodies is getting smaller $U$ is getting larger.
When all the bodies have vanished completely the flow rate $U$ is equal to 1.
In Fig.~\ref{fig:Eroding20flowrate}, the flow rate increases exponentially 
with growth rate close to $\ln10^4$. }

\begin{figure}[H]
\center
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/flow_rate20dense}
%\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20flowrate} The incoming flow rate of $U$ in log scale versus time. 
Under fixed pressure drop condition, $U$ will be small initially,
increase when bodies are eroding, and grow exponentially 
to 1 when there is no body in the fluid. }
\end{figure}

Snapshots of the
bodies at four equispaced times are shown in
Figure~\ref{fig:Eroding20vort}.   The colormap is the vorticity which is
equivalent to the shear stress when restricted to the eroding bodies.
Therefore, erosion is fastest in regions where the magnitude of the
vorticity is largest.

@@

\begin{figure}[H]
\begin{center}
  \includegraphics[height=0.23\textwidth]{./figs/20b_dense1}
  \includegraphics[height=0.23\textwidth]{./figs/20b_dense101}
  \includegraphics[height=0.23\textwidth]{./figs/20b_dense201}
  \includegraphics[height=0.23\textwidth]{./figs/20b_dense301}
%\includegraphics[width = 0.42 \textwidth]{./figs/20b_dense1}
%\includegraphics[width = 0.507 \textwidth]{./figs/20b_dense101}\\
%\includegraphics[width = 0.42 \textwidth]{./figs/20b_dense201}
%\includegraphics[width = 0.507 \textwidth]{./figs/20b_dense301}
\caption{\label{fig:Eroding20vort} Erosion of 20 nearly touching bodies
  with the fixed pressure drop condition. The color is the vorticity of
  the fluid. The 4 snapshots are evenly spaced in time. We use $N_\iin =
  256$ discretization points on the body, a time-step of $\Dt=10^{-4}$,
  and smoothing parameters of $\eps= 15/256$ and $\sigma=10/256 $. In
  the forth frame, bodies 4, 12, and 17 have vanished completely, and
  body 19 has almost completely eroded.}
\end{center}
\end{figure}

In this example, we also observe that erosion creates a network of
channels from the inlet to outlet where the velocity and vorticity, and
therefore erosion rate, are much larger than the other regions.  These
channels can be further visualized by considering the streamlines at
several configurations of the eroding body.  In
Figure~\ref{fig:Eroding20tracer}, we simulate 200 tracers that are
initially to the left of the porous region and equispaced in $y$.  There
are two clear region where the tracers pass through the geometry
much faster than the other regions.  To illustrate the ability of the
Barycentric quadrature formula, streamlines very close to the eroding
bodies are illustrated in the last frame of Figure~\ref{fig:Eroding20tracer}.

\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150}\\

\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270_zoom}\\

%\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210}
%\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b240}
%\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270}
\caption{\label{fig:Eroding20tracer}The trajectories of 200 tracers in
  the fluid with 20 bodies. The five first snapshots are evenly spaced in time.
  The tracers are evenly spread in the $y$-direction and released at
  $x=-1$. The sixth frame is a zoom window of the fifth frames. 
Since we use a Barycentric quadrature rule
  and the fourth order Runge Kutta method, the tracers can be very close
  to but not penetrate the bodies.}
\end{center}
\end{figure}

@@

Since we use a Barycentric quadrature rule to approximate the velocity
of tracers and apply the fourth order Runge Kutta method to find the new
locations of tracers in each step, the tracers can be very close to but
not penetrate the bodies.  Those body-closed tracers are moving
perfectly so we are able to observe the details of the interaction of
bodies and fluid.  Although the direction of flow is from left to right,
the tracers will not only travel horizontally but also vertically.  In
the fifth frame of Fig.~\ref{fig:Eroding20tracer}, some tracers
climb up from right to left around the body.  There is a zoom windows of
the climbing tracers trajectories in the last frame of Fig.~\ref{fig:Eroding20tracer} and
they are taken from the fifth frame of
Fig.~\ref{fig:Eroding20tracer} with more tracers.  In the last frame of 
Fig.~\ref{fig:Eroding20tracer} the tracers around the center body first go
from the channel below the body, climb up to the left, then go to the
right and join another tracers in the channel above the body. 
}
%\begin{figure}[H]
%\begin{center}
%\includegraphics[width = 0.32\textwidth]{./figs/tracer_20b210_zoom}
%\includegraphics[width = 0.32\textwidth]{./figs/tracer_20b240_zoom}
%\includegraphics[width = 0.32\textwidth]{./figs/tracer_20b270_zoom}
%\caption{\label{fig:Eroding20zoom}The zoom windows in the last three
%  frames of Figure~\ref{fig:Eroding20tracer}. The trajectories of
%  tracers around the bodies. Since we use a Barycentric quadrature rule
 % and the fourth order Runge Kutta method, the tracers can be very close
 % to but not penetrate the bodies.}
%\end{center}
%\end{figure}
{\color{red}

@@

{\color{red}
In Fig.~\ref{fig:Eroding20tort}(c) we calculate the difference of 
$\tau$ at two neighboring tracers 
and draw the trajectories of neighboring tracers
if the difference of $\tau$ at the neighboring tracers is large. 
From Fig.~\ref{fig:Eroding20tort}(c),
 we are able to tell the neighboring trajectories will be very 
different if they are separated by bodies. In Fig.~\ref{fig:Eroding20tort}(c)
we collect the 10 largest difference of $\tau$ at the neighboring tracers and 
draw the pair of trajectories in the same color. For each pair of trajectories, 
they are splitted by the body at the left side of the body, then
one of them goes up and the other one goes down along the body 
until they rejoin with each other at the right side of body.  
Furthermore, the pair trajectories could be splitted by more than one body, 
for example in Fig.~\ref{fig:Eroding20tort}(c) the toppest red pair trajectories 
are splitted by two close bodies.
}

\begin{figure}[H]
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc20_268}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_local20_268}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_diff_top10_268}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20tort} The tortuosity study in the fluid with 
20 nearly touching bodies when the porosity is 62.9 \%.
(a) The 
x-component velocity u(-1, y) with respect to its maximum velocity $u_{max}=2.9828
\times 10^{-3}$ at equal spreading tracers on the cross section x = -1. 
(b) The local tortuosity $\tau(y)$ on the cross section x = -1. 
(c) The largest 10 difference of 
local tortuosity $\tau$ at two neighboring tracers in
 20 nearly touching bodies eroding when the porosity is 62.9 \% 
and the tortuosity is 1.0979. The same color of a pair of trajectories 
means they are from two neighboring tracers.}
%(c) The difference of $\tau(y)$ at two neighboring tracers on the cross section x = -1.
%(d) The ratio of the minimum to maximum speeds along the trajectories of tracers.}
\end{figure}
{\color{red}
In Fig.~\ref{fig:Eroding20area} we
measure the area occupied by bodies 
during the eroding. In the beginning, 62.32\% of area of $[-1,1] \times
[-1,1]$ is occupied by 20 bodies. Then the eroding speed is increasing
when more and more bodies are vanished in the fluid and 
the maximum eroding speed happens when 7.57\% of area of $[-1,1] \times
[-1,1]$ is occupied by 8 bodies. After that the speed drops rapidly to 0 when
all the bodies are vanished completely.}

\begin{figure}[H]
\center
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[width =\linewidth]{./figs/porosity20dense}
%\caption{}
\end{subfigure}%
\caption{\label{fig:Eroding20area}  The area fraction of 20 nearly
  touching eroding bodies and the eroding speed versus time. 
}
\end{figure}

%\begin{figure}[H]
%\center
%\includegraphics*[width =0.5\linewidth]{./figs/tort_diff_top10_268}
%\caption{\label{fig:Eroding20tort_traj} The largest 10 difference of 
%local tortuosity $\tau$ at two neighboring tracers in
% 20 nearly touching bodies eroding when the porosity is 62.9 \% 
%and the tortuosity is 1.0979. The same color of a pair of trajectories 
%means they are from two neighboring tracers.}
%\end{figure}

@@

\begin{figure}[H]
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_local100}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width = \textwidth]{./figs/100b_t100tracer}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100tort} The tortuosity study in the fluid with 100 bodies 
when the porosity is 62.98 \%.
(a) The 
x-component velocity u(-1, y) with respect to its maximum velocity $u_{max}=3.9021
\times 10^{-4}$ at equal spreading tracers on the cross section x = -1. 
(b) The local tortuosity $\tau(y)$ on the cross section x = -1. 
(c) The historical trajectories of 
200 tracers released at x = -1 in the fluid with multiple bodies.}
%(c) The difference of $\tau(y)$ at two neighboring tracers on the cross section x = -1.
%(d) The ratio of the minimum to maximum speeds along the trajectories of tracers.}
\end{figure}
